---
title: "CF duration analysis - whole audio version"
author: "Thejasvi Beleyur"
date: "7/6/2021"
output: html_document
---

This is a follow up analysis to delve deeper into the apparent decrease in CF duration 

This notebook will detail the use of bayesian linear models and mixed models to 
quantify the difference between single and multi-bat call CF duration only. 

The final results are presented in the [Final Summary](#finalsummary).

All intervals reported here are 93% highest posterior density compatibility intervals [@gelman2019confidence;@mcelreath2020statistical]. 

```{r loading, echo=T, warning=F,message=F,results='hide'}

library(arm)
library(coda)
library(lme4)
library(stringi)
library(stringr)
set.seed(82319)
d <- read.csv('../verified_cf_detections.csv')
# create a single/multi-bat categorical variable
d['multi_bat'] <- factor(d['num_bats']>1)
d['annot_id'] <- stri_sub(d$audio_file,28,-4)
d['annot_id'] <- str_replace_all(d$annot_id, '21502300','2123')
# sort the dataset by time first
time_ordering <- stri_order(d$annot_id,numeric=TRUE)
d_tordered <- d[time_ordering,]
# - is there an influence of the recording hour? 
split.and.join <- function(x){
  splitx <- str_split(x,'_')
  jointx <- paste(splitx[[1]][1],'_',splitx[[1]][2],sep='')
  jointx}
d_tordered$rechour <- sapply(d_tordered$annot_id,split.and.join)

# also scale up some of the measurements into more meaningful units
# -- here we'll multiply by 10^3 to convert s to ms
d_tordered$duration <- d_tordered$duration*10^3

num_post <- 5000
# convenience function 
get.hpd <- function(x,prange=0.95){HPDinterval(as.mcmc(x),prob=prange)}

get.map.coi <- function(x){c(mean(x), get.hpd(x))}


```
## Quick data plotting 

Let's plot our data first to see if everything about it is okay. 

```{r quickplot}
boxplot(duration~as.factor(multi_bat), data=d_tordered, ylab='Duration, ms', xlab='Multi-bat or not')

```
Just looking at the data


## Linear model: starting simple 


```{r lm-model, warning=F}
d_tordered$invdurn <- d_tordered$duration^-1
lm0 <- lm(invdurn~as.factor(multi_bat), data=d_tordered)


par(mfrow=c(3,2))
plot(lm0)
acf(resid(lm0))

```

The 1/y transform seems to work best in making the model fit the normal assumptions. I also tried the square-root and log10 transforms - but they didn't work as well. 

We see there is some autocorrelation in the residuals - which is kind of expected since when there are 
multiple calls from the same audio file, we can expect that the CF durations are also similar. This tells us 
we need to include annotation ID as a factor in a mixed effects model. 

## Linear mixed-model: accounting for correlations within each audio clip
Let's build a LMM with the annotation-ID (recording ID) and recording hour as the random intercepts, and
the group size (single/multi) as a fixed effect. There can be multiple CF duration measurements 
from the same recording, and the same bats may be seen within the span of an hour, so this is likely to be the main source of correlation in the dataset.

*Note*: ```rechour``` added as a random intercept on 2021-07-30, and the  ```rechour``` as a fixed effect analysis aren't reported in the paper anymore. 

```{r lmm-model}
lmm0 <- lmer(invdurn~as.factor(multi_bat)+(1|annot_id) +(1|rechour), data=d_tordered)


par(mfrow=c(2,3))
# check residuals vs multi_bat
plot(d_tordered$multi_bat,resid(lmm0),main='residuals - duration,1/ms') 
# do residuals match the normal distbn assumption - qqplot
qqnorm(resid(lmm0));qqline(resid(lmm0))
# do random effects match normal distbn assumption - qqplot
qqnorm(ranef(lmm0)$annot_id$`(Intercept)`);qqline(ranef(lmm0)$annot_id$`(Intercept)`,main='random effects') 
# do random effects match normal distbn assumption - qqplot
qqnorm(ranef(lmm0)$rechour$`(Intercept)`);qqline(ranef(lmm0)$rechour$`(Intercept)`,main='random effects') 
# check crudely for autocorrelation of residuals 
acf(resid(lmm0))
```

The assumptions of the LMM seem to be met fairly okay - though not perfectly. Now, let's proceed to run simulations to and estimate parameter values. 

```{r param-estimation}

cfdurn.sim <- sim(lmm0, n.sim=num_post)
cfdurn.fixefs.hpd <- round(apply(fixef(cfdurn.sim),2,get.hpd),2)
single_cfdurn <- 1/fixef(cfdurn.sim)[,1]
multi_cfdurn <- 1/(fixef(cfdurn.sim)[,1]+fixef(cfdurn.sim)[,2])

# difference in ms:
cfdurn.diff <- multi_cfdurn - single_cfdurn
cfdurn.diff.mapcoi <- get.map.coi(cfdurn.diff)

# Mean MAP and CoI
cfdurn.single.mapcoi <- get.map.coi(single_cfdurn)
cfdurn.multi.mapcoi <- get.map.coi(multi_cfdurn)

cfdurn.rel.increase <- multi_cfdurn/single_cfdurn
# ratio MAP and CoI
cfdurn.rel.mapcoi <- get.map.coi(cfdurn.rel.increase)


single.cfdurn.hpd <- round(get.hpd(single_cfdurn),2)
multi.cfdurn.hpd <- round(get.hpd(multi_cfdurn),2)
cfdurn.rel.hpd <- round(get.hpd(cfdurn.rel.increase),2)

```


### Round 2 analysis (2021-07-14) (not used anymore)
These are my attempts at checking to see if the small peak in the residual autocorrelation is perhaps because I didn't consider an additional variable (eg. recording hour). Even though I don't have reason to, I also see that the recording hour itself may have systematic effects because the same bat flew repeatedly within a period of time. Broadly speaking, the 'null' result from the primary analysis is not surprising.

```{r post-analysis}
# lmm.post <- lmer(invdurn~as.factor(multi_bat)+(1|annot_id)+rechour, data=d_tordered)
# 
# 
# par(mfrow=c(3,2))
# # check residuals vs multi_bat
# plot(d_tordered$multi_bat,resid(lmm.post),main='residuals - duration,1/ms') 
# # do residuals match the normal distbn assumption - qqplot
# qqnorm(resid(lmm.post));qqline(resid(lmm.post))
# # do random effects match normal distbn assumption - qqplot
# qqnorm(ranef(lmm.post)$annot_id$`(Intercept)`);qqline(ranef(lmm.post)$annot_id$`(Intercept)`,main='random effects') 
# # check crudely for autocorrelation of residuals 
# acf(resid(lmm.post))
# # check if the residuals are similar across recording hours 
# boxplot(resid(lmm.post)~d_tordered$rechour)
```

The quick-check inclusion of the recording hour as a factor also doesn't seem to make the small autocorrelation peak go away. I looked at the residuals themselves once more - there doesn't seem to be a 'trend'. Despite including annotation ID as a random intercept, I guess there's only so much that even the LMM can handle. 

```{r lmmpost}
# 
# 
# cfdurn.sim.post <- sim(lmm.post, n.sim=num_post)
# cfdurn.fixefs.hpd.post <- round(apply(fixef(cfdurn.sim.post),2,get.hpd),2)
# single_cfdurn.post <- 1/fixef(cfdurn.sim.post)[,1]
# multi_cfdurn.post <- 1/(fixef(cfdurn.sim.post)[,1]+fixef(cfdurn.sim.post)[,2])
# 
# cfdurn.rel.increase.post <- multi_cfdurn.post/single_cfdurn.post
# 
# 
# single.cfdurn.hpd.post <- round(get.hpd(single_cfdurn.post),2)
# multi.cfdurn.hpd.post <- round(get.hpd(multi_cfdurn.post),2)
# cfdurn.rel.hpd.post <- round(get.hpd(cfdurn.rel.increase.post),2)

```

We see the relative change in CF duration still remains very similar (close to the 0.9-1.1 range). The main difference is in the increase in highest posterior density interval of the single bat CF duration. Now both intervals are very overlapping. It seems like including recording hour helped in explaining some systematic differences that were there across time. 


### Final summary {#finalsummary}

The final results do not seem to support a decrease in cF duration in groups, as suggested by the individual call analysis. The 
multi-call CF duration analysis is more robust because of the increased sample sizes - and so I'd side more with these results.



#### Model with  group size and recording hour as fixed effect, and annotation id as random intercept



```{r write-results}

df <- rbind(cfdurn.diff.mapcoi, cfdurn.rel.mapcoi, cfdurn.single.mapcoi, cfdurn.multi.mapcoi)
row.names(df) <- c('durn-difference','durn-ratio','single','multi')
colnames(df)<- c('MAP','lower-95pcHPD', 'upper-95pcHPD')
write.csv(df, 'cf_duration_change_summary.csv')
```



```{r sessioninfo}
sessionInfo()
```
